---
title: "Beaufort Report Appendices"
author: "Jonah Bacon"
date: "October 01, 2022"
output: word_document
---

This document produces the Appendices for the Beaufort Report then outputs it to MS Word. 
Copy these appendices from the Word doc into Excel for final formatting to remain consistent. 

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(knitr)
library(lubridate)

curr_yr <- 2022
load(here::here("../../Data/Database/PrudhoeCatch&LengthDataset_2001-2022_Version5.Rdata"))


#First the Data Prep

#### EFFORT
#source("Effort&CPUE.R")

# Could call Effort&CPUE.R file but it's slow and fills environment with a lot of files
# So instead I copied the only relevant part here
all.len <- all.len %>% mutate(Net = paste0(Station, Side))

effort <- full_join(all.len %>%
                    distinct(EndDate, Net), 
                    allcatch %>% distinct(EndDate, Net), 
          by = c("EndDate", "Net")) %>% 
  left_join(all.len %>%
            dplyr::select(EndDate, Net, StartDateTime, EndDateTime), 
            by = c("EndDate", "Net")) %>% 
  distinct(EndDate, Net, StartDateTime, EndDateTime) %>%
  #this first mutate MANUALLY adds in skipped dates
  mutate(
    # Add start/end date/time for fishing when we caught fish but didn't have any lengths:
        StartDateTime=replace(StartDateTime, EndDate=="2022-07-16" & Net == "220E", 
                               as.POSIXct("2022-07-15 07:20")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-07-16" & Net == "220E", 
                             as.POSIXct("2022-07-16 6:57")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-07-31" & Net == "220E", 
                               as.POSIXct("2022-07-30 07:22")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-07-31" & Net == "220E", 
                             as.POSIXct("2022-07-31 7:30")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-02" & Net == "220E", 
                               as.POSIXct("2022-08-01 06:36")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-02" & Net == "220E", 
                             as.POSIXct("2022-08-02 06:48")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-05" & Net == "220W", 
                               as.POSIXct("2022-08-04 6:20")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-05" & Net == "220W", 
                             as.POSIXct("2022-08-05 6:37")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-10" & Net == "220W", 
                               as.POSIXct("2022-08-09 06:54")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-10" & Net == "220W", 
                             as.POSIXct("2022-08-10 06:17")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-18" & Net == "220W", 
                               as.POSIXct("2022-08-17 07:54")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-18" & Net == "220W", 
                             as.POSIXct("2022-08-18 06:50")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-20" & Net == "220E", 
                               as.POSIXct("2022-08-19 06:10")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-20" & Net == "220E", 
                             as.POSIXct("2022-08-20 06:57"))) %>%
  
  # Add start date/times for fishing when we caught 0 fish:
    add_row(
    EndDate = as.Date("2022-07-30"),
    Net = "220E",
    StartDateTime=as.POSIXct("2022-07-29 7:32"),
    EndDateTime = as.POSIXct("2022-07-30 07:22")) %>%

  add_row(
    EndDate = as.Date("2022-08-06"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-05 06:44"),
    EndDateTime = as.POSIXct("2022-08-06 07:41")) %>%

  add_row(
    EndDate = as.Date("2022-08-07"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-06 07:41"),
    EndDateTime = as.POSIXct("2022-08-07 06:23")) %>%

  add_row(
    EndDate = as.Date("2022-08-15"),
    Net = "220W",
    StartDateTime = as.POSIXct("2022-08-14 06:37"),
    EndDateTime = as.POSIXct("2022-08-15 06:42")) %>%

  add_row(
    EndDate = as.Date("2022-08-15"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-14 06:46"),
    EndDateTime = as.POSIXct("2022-08-15 06:52")) %>%

  add_row(
    EndDate = as.Date("2022-08-18"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-17 08:07"),
    EndDateTime = as.POSIXct("2022-08-18 06:55")) %>%

  mutate(Year = year(EndDate),
         Effort = as.numeric(EndDateTime - StartDateTime)) %>%
  arrange(EndDate, Net)

# Now summarize the season
startsampling <- all.len %>% filter(Year == curr_yr) %>% summarise(date(min(StartDateTime))) %>% as.vector()
endsampling <- all.len %>% filter(Year == curr_yr) %>% summarise(date(max(EndDateTime)))

# Make a list of every date and station combo, from first date to last
samplesumm <- expand.grid(EndDate = seq.Date(from=startsampling[1,1], to=endsampling[1,1], by = "day"), 
            Net = allcatch %>% filter(Year == curr_yr) %>% distinct(Net) %>% unlist()) %>%
  mutate(Year = year(EndDate)) %>%
  #now add to this with the number of net nights
  left_join(effort %>% filter(Year == curr_yr) %>% 
              mutate(setdays = as.numeric(date(EndDateTime)- date(StartDateTime)),
                     Station = substr(Net, 1,3)) %>% dplyr::select(Year, EndDate, Net, setdays),
            by = c("Year" = "Year", "EndDate" = "EndDate", "Net" = "Net"))


# Make summary function for when we summarize by all LGs combined
summ_AllLG <- function(species) {
  .tempdf <- samplesumm %>% left_join(allcatch %>% filter(Year == curr_yr, Species == species) %>% 
                         #have to group all LGs together and sum, so following line is diff from indiv LG summaries
                         group_by(EndDate, Net) %>% summarise(totcount = sum(totcount)) %>%
                         dplyr::select(EndDate, Net, totcount), by = c("EndDate" = "EndDate", "Net" = "Net")) %>%
    replace_na(list(totcount = 0)) %>% #IMPORTANT! This differentiates btwn no catch and no sampling!
    mutate(CPUE = totcount / setdays) %>% dplyr::select(-Year, -totcount, -setdays) %>%
    spread(Net, value = CPUE)
  kable(.tempdf, digits = c(0))
}

# Make summary function for when we summarize by each LG
summ_ByLG <- function(species, LengthGroup) {
  .tempdf <- samplesumm %>% left_join(allcatch %>% filter(Year == curr_yr, Species == species, LG == LengthGroup) %>%
                         dplyr::select(EndDate, Net, totcount), by = c("EndDate" = "EndDate", "Net" = "Net")) %>%
  replace_na(list(totcount = 0)) %>% #IMPORTANT! This differentiates btwn no catch and no sampling!
  mutate(CPUE = totcount / setdays) %>% dplyr::select(-Year, -totcount, -setdays) %>%
  spread(Net, value = CPUE) 
  
  kable(.tempdf, digits = c(0))
}


```



# Arctic Cisco
## ARCS All Length Groups
```{r Appx3, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("ARCS")
```


## ARCS LG1
```{r Appx4, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("ARCS", 1)
# If you have any errors in here that say "Error: DUplicate identifiers for rows (326, 327)"
# then you have two or more totals for the same LG! Fix this
```


## ARCS LG2
```{r Appx5, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("ARCS", 2)

```



## ARCS LG3
```{r Appx6, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("ARCS", 3)

```


# Broad Whitefish
## BDWF All Length Groups
```{r Appx7, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("BDWF")
```


## BDWF LG1
```{r Appx8, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("BDWF", 1)

```



## BDWF LG2
```{r Appx9, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("BDWF", 2)
```



## BDWF LG3
```{r Appx10, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("BDWF", 3)
```

---

# Least Cisco
## LSCS All Length Groups
```{r Appx11, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("LSCS")
```

---

## LSCS LG1
```{r Appx12, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("LSCS", 1)
```

---

## LSCS LG2
```{r Appx13, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("LSCS", 2)
```

---


# Dolly Varden
## DLVN All Length Groups
```{r Appx14, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("DLVN")
```


## DLVN LG1
```{r Appx15, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("DLVN", 1)
```

---


## DLVN LG2
```{r Appx16, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_ByLG("DLVN", 2)
```

---

# Humpback Whitefish
## HBWF All Fish
```{r Appx17, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("HBWF")
```

---

# Arctic Cod
## ARCD All Fish
```{r Appx18, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("ARCD")
```

---

# Saffron Cod
## SFCD All Fish
```{r Appx19, echo = FALSE}
options(knitr.kable.NA = ' ')

summ_AllLG("SFCD")
```

