---
title: "Beaufort Report Tables"
author: "Jonah Bacon"
date: "Oct 01, 2022"
output: word_document
---
This document produces tables for the Beaufort Report then outputs it to MS Word. 
The tables from the Word doc can then be copied into Excel for final formatting to remain consistent. 
If you are looking at this in RStudio, press the "Knit" button to create the Word document.

```{r setup, include=FALSE}
library(tidyr)
library(dplyr)
library(knitr)
library(lubridate)
#library(kableExtra) #great package for formatting kable in pdf/html but won't work with word
library(pander)
library(vegan)
#library(flextable)

knitr::opts_chunk$set(echo = TRUE)
load(here::here("../../Data/Database/PrudhoeCatch&LengthDataset_2001-2022_Version5.Rdata"))
allcatch$Station <- substr(allcatch$Net, 1, 3)

curr_yr <- 2022
spptot_table <- allcatch %>% filter(Year == curr_yr) %>% group_by(Species, Station) %>%
  summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>%
  spread(Station, TotalCatch) %>% ungroup() %>% 
  mutate('Total Catch' = rowSums(.[2:5], na.rm = TRUE)) %>%
  left_join(spplookup %>% dplyr::select(-familygroup), 
            by = c("Species" = "Species"))%>%
  dplyr::select(-Species) %>% rename(Species = commonname) %>%
  mutate(Species = as.character(Species)) %>% dplyr::select(Species, everything()) %>%
  bind_rows(colSums(.[-1], na.rm = TRUE)) %>%
  replace_na(list(Species = "Total Catch"))

# need to fix station order to 220 -> 218 -> 214 -> 230


numspp <- spptot_table %>% filter(Species != "Hybrid Cisco spp.") %>% 
  summarize(length(unique(Species))) %>% pull() #last part turns back into vector


# MAKE SURE YOU EXCLUDE HYBRIDS AND UNKNOWNS
allsitesdiv <- allcatch %>% filter(Year == curr_yr, Species != "HYCS") %>% group_by(Species) %>%
  summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>%
  summarise(ntot = sum(TotalCatch>0), stntot = sum(TotalCatch, na.rm = TRUE),
            shan = diversity(TotalCatch, index="shannon"),
            even = shan/log(numspp)) 



diversity <- allcatch %>% filter(Year == curr_yr, Species != "HYCS") %>% 
  group_by(Species, Station) %>%
  summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>% group_by(Station) %>%
  summarise(ntot = sum(TotalCatch>0), stntot = sum(TotalCatch, na.rm = TRUE),
            shan = diversity(TotalCatch, index="shannon"),
            even = shan/log(numspp)) %>% bind_rows(allsitesdiv) %>% t() %>% data.frame() 




#### EFFORT
#source("Effort&CPUE.R")

# Could call Effort&CPUE.R file but it's slow and fills environment with a lot of files
# So instead I copied the only relevant part here
all.len <- all.len %>% mutate(Net = paste0(Station, Side))

effort <- full_join(all.len %>%
                    distinct(EndDate, Net), 
                    allcatch %>% distinct(EndDate, Net), 
          by = c("EndDate", "Net")) %>% 
  left_join(all.len %>%
            dplyr::select(EndDate, Net, StartDateTime, EndDateTime), 
            by = c("EndDate", "Net")) %>% 
  distinct(EndDate, Net, StartDateTime, EndDateTime) %>%
  #this first mutate MANUALLY adds in skipped dates
  mutate(
    # Add start/end date/time for fishing when we caught fish but didn't have any lengths:
        StartDateTime=replace(StartDateTime, EndDate=="2022-07-16" & Net == "220E", 
                               as.POSIXct("2022-07-15 07:20")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-07-16" & Net == "220E", 
                             as.POSIXct("2022-07-16 6:57")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-07-31" & Net == "220E", 
                               as.POSIXct("2022-07-30 07:22")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-07-31" & Net == "220E", 
                             as.POSIXct("2022-07-31 7:30")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-02" & Net == "220E", 
                               as.POSIXct("2022-08-01 06:36")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-02" & Net == "220E", 
                             as.POSIXct("2022-08-02 06:48")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-05" & Net == "220W", 
                               as.POSIXct("2022-08-04 6:20")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-05" & Net == "220W", 
                             as.POSIXct("2022-08-05 6:37")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-10" & Net == "220W", 
                               as.POSIXct("2022-08-09 06:54")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-10" & Net == "220W", 
                             as.POSIXct("2022-08-10 06:17")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-18" & Net == "220W", 
                               as.POSIXct("2022-08-17 07:54")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-18" & Net == "220W", 
                             as.POSIXct("2022-08-18 06:50")),
         
         StartDateTime=replace(StartDateTime, EndDate=="2022-08-20" & Net == "220E", 
                               as.POSIXct("2022-08-19 06:10")), 
         EndDateTime=replace(EndDateTime, EndDate=="2022-08-20" & Net == "220E", 
                             as.POSIXct("2022-08-20 06:57"))) %>%
  
  # Add start date/times for fishing when we caught 0 fish:
    add_row(
    EndDate = as.Date("2022-07-30"),
    Net = "220E",
    StartDateTime=as.POSIXct("2022-07-29 7:32"),
    EndDateTime = as.POSIXct("2022-07-30 07:22")) %>%

  add_row(
    EndDate = as.Date("2022-08-06"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-05 06:44"),
    EndDateTime = as.POSIXct("2022-08-06 07:41")) %>%

  add_row(
    EndDate = as.Date("2022-08-07"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-06 07:41"),
    EndDateTime = as.POSIXct("2022-08-07 06:23")) %>%

  add_row(
    EndDate = as.Date("2022-08-15"),
    Net = "220W",
    StartDateTime = as.POSIXct("2022-08-14 06:37"),
    EndDateTime = as.POSIXct("2022-08-15 06:42")) %>%

  add_row(
    EndDate = as.Date("2022-08-15"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-14 06:46"),
    EndDateTime = as.POSIXct("2022-08-15 06:52")) %>%

  add_row(
    EndDate = as.Date("2022-08-18"),
    Net = "220E",
    StartDateTime = as.POSIXct("2022-08-17 08:07"),
    EndDateTime = as.POSIXct("2022-08-18 06:55")) %>%

  mutate(Year = year(EndDate),
         Effort = as.numeric(EndDateTime - StartDateTime)) %>%
  arrange(EndDate, Net)

# A few NAs occur when we have catch but no lengths for that day 
# (because the times are recorded in the length dataframe!)
# So run this and manually check for any NAs in the current year

```


First is Table 3, which is actually a combination of two tables: catch totals and a diversity summary.   
```{r table3.1, echo=FALSE, message=FALSE}
#most formatting is done using package kableextra but this DOESN'T work with .DOCX
options(knitr.kable.NA = '-')
spptot_table %>% knitr::kable(format.args = list(big.mark = ","))

```

\newpage 
\pagebreak
The following is the same table but formatted using a different package, and this time the diversity table follows.  

```{r table3.2, echo=FALSE}
#pagebreaks and newpage dont work for word. If you need this, add package "worded"


#set.alignment('left', row.names = 'right')
spptot_table %>% pander::pander(missing = "-", big.mark = ",", 
                                justify = c('left', 'center', 'center','center','center', "right"))

```
\pagebreak

```{r divertable.3, echo=FALSE, message=FALSE}

#requires "Pandoc" installed, at least V2
#https://pandoc.org/installing.html make sure you restart comp after install

#spptot_table %>% regulartable() #looks like crap in default

diversity %>% slice(c(2, 4, 5)) %>% pander()

```

```{r Table1, echo = FALSE}



effort %>% filter(Year == curr_yr & Effort != is.na(Effort)) %>% 
  mutate(Station = substr(Net, 1,3)) %>%
  group_by(Station) %>% summarise(Start_Date = min(date(StartDateTime)),
                                  End_Date = max(EndDate)) %>%
  #now add Net Nights
  left_join(effort %>% filter(Year == curr_yr) %>% 
  mutate(setdays = as.numeric(date(EndDateTime)- date(StartDateTime)),
         Station = substr(Net, 1,3)) %>%
  group_by(Net, Station) %>% summarise(totalsetdays=sum(setdays)) %>%
  group_by(Station) %>% summarise(NetNights = max(totalsetdays)),
  by = c("Station" = "Station")) %>% kable()

```

The following two tables (Table 2) will need to be united together in Excel
```{r Table2, echo = FALSE}
env_allyears %>% dplyr::select(-c(Temp_Top, Temp_Mid, Temp_Bot, Temp_Bot_1.5)) %>%
  gather(measurement, ppt, -Year, -Date, -Station) %>%
  group_by(Year) %>% 
  summarise(MeanSalin=mean(ppt, na.rm = TRUE), SD=sd(ppt, na.rm = TRUE), 
            Min=min(ppt, na.rm = TRUE), Max=max(ppt, na.rm = TRUE)) %>%
  mutate(Range = paste0(Min, "–", Max)) %>%
  dplyr::select(Year, MeanSalin, Range, SD) %>%
  kable(digits = c(0, 4, 0, 4))


env_allyears %>% dplyr::select(-c(Salin_Top, Salin_Mid, Salin_Bot, Salin_Bot_1.5)) %>% 
  gather(measurement, temp, -Year, -Date, -Station) %>%
  group_by(Year) %>% 
  summarise(MeanTemp=mean(temp, na.rm = TRUE), SD=sd(temp, na.rm = TRUE), 
            Min=min(temp, na.rm = TRUE), Max=max(temp, na.rm = TRUE)) %>%
  mutate(Range = paste0(Min, "–", Max)) %>%
  dplyr::select(Year, MeanTemp, Range, SD) %>%
  kable(digits = c(0, 4, 0, 4))

```

Now Table 4:
```{r Table4, echo=FALSE, message=FALSE, warning=FALSE}
all.len %>% left_join(spplookup %>% dplyr::select(-familygroup), 
            by = c("Species" = "Species")) %>%
  dplyr::select(-Species) %>% rename(Species = commonname) %>%
  filter(Year == curr_yr) %>% group_by(Species) %>% 
  summarise(NumberSampled = n(), MeanLengthmm = mean(Length), MinLength = min(Length), MaxLength = max(Length), StandardDeviation = sd(Length)) %>%
  mutate(LengthRangemm = paste0(MinLength, "–", MaxLength)) %>%
  dplyr::select(Species, NumberSampled, MeanLengthmm, LengthRangemm, StandardDeviation) %>%
  kable(format.args = list(big.mark = ","), digits = c(0, 0, 1, 0, 4))


```
