---
title: "Beaufort Report Summary of 2022 Season"
author: "Jonah Bacon"
date: "October 01, 2022"
output:
  html_document:
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
  
---


```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
# If you get a line 16 error, check that all "ggsave" lines in effort file are commented off 
# You don't want to save files, you just want to import them into here. 

library(tidyverse)
library(knitr)
library(kableExtra)
library(here)

curr_yr <- 2022   #adjust this for the year of interest
load(here::here("../../Data/Database/PrudhoeCatch&LengthDataset_2001-2022_Version5.Rdata"))
source(here::here("Effort&CPUE.R"))


# Subset for current year specified above
allcatch.year <- allcatch %>% dplyr::filter(Year == curr_yr) 
all.len.year  <- all.len  %>% dplyr::filter(Year == curr_yr)
effort.year   <- effort   %>% dplyr::filter(Year == curr_yr)

# Add Station column to all for ease later. Rename to LengthGroup for clarity
allcatch.year <- allcatch.year %>% mutate(Station = substr(Net, 1, 3)) %>% 
  rename(LengthGroup = LG)
all.len.year  <- all.len.year  %>% mutate(Station = substr(Net, 1, 3)) %>% 
  rename(LengthGroup = Group)
effort.year   <- effort.year %>% mutate(Station = substr(Net, 1, 3)) 

```

# Season Effort

The total number of hours fished by each site is first divided by 24 to convert hours to days, then divided by 2.
This converts cod-end nights (2 per station) to station nights (station nights = net nights; for our purposes here they are the same).
A 'Net Night' is defined as the number of 24 hour periods fished across the entire season.
One net night = 24 hours of fishing by any station.
For example, if Station 230 fishes for 24 hours, that site produces 1 net night in that 24 hour period.


```{r effort_summ, include = FALSE}

# Net nights per station. Check outputs for realistic number of days fished.
Eff214 <- effort.year %>% filter(Station == "214") %>% 
  summarise(sum(as.numeric(Effort), na.rm=TRUE)/48) %>% as.double() %>% round(digits = 1)
Eff218 <- effort.year %>% filter(Station == "218") %>% 
  summarise(sum(as.numeric(Effort), na.rm=TRUE)/48) %>% as.double() %>% round(digits = 1)
Eff220 <- effort.year %>% filter(Station == "220") %>% 
  summarise(sum(as.numeric(Effort), na.rm=TRUE)/48) %>% as.double() %>% round(digits = 1)
Eff230 <- effort.year %>% filter(Station == "230") %>% 
  summarise(sum(as.numeric(Effort), na.rm=TRUE)/48) %>% as.double() %>% round(digits = 1)

totaleffort <- sum(Eff214, Eff218, Eff220, Eff230)
totaleffort.df <- data.frame(Station = c("214", "218", "220", "230"), 
                             Effort_netnights = c(Eff214, Eff218, Eff220, Eff230))

```
In `r curr_yr`, there were `r totaleffort` "net nights" among all four sampling stations.  
At Station 230, there were `r Eff230` net nights, `r Eff214` net nights at Station 214, 
`r Eff218` net nights at Station 218, and `r Eff218` net nights at Station 218. 

```{r efforttable, echo = FALSE}
kable(totaleffort.df) %>%  
        kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) 
```


# Catch Summary - All Species

```{r catchsummary, echo = FALSE}
catchbyyear <- allcatch %>% group_by(Year, Species) %>% summarise(anncount = sum(totcount)) %>%
  spread(Species, value = anncount) %>% replace(., is.na(.), 0) %>% ungroup() %>%
  mutate(Year = as.character(Year)) # this is to prevent thousands comma in the year

kable(catchbyyear %>%  dplyr::select(Year, ARCD:GRAY), format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) 

kable(catchbyyear %>%  dplyr::select(Year, HBWF:RKGL), format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) 

kable(catchbyyear %>%  dplyr::select(Year, SFCD:WSGL), format.args = list(big.mark = ",")) %>%
  kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) 

```


# Individual Species Summaries

```{r spp_fxn, include = FALSE, warning = FALSE, results = 'asis'}

spp_summ <- function(x){
  # first, filter by current species
  .spp_len <- all.len.year %>% filter(Species == x)
  .spp_tot <- allcatch.year %>% filter(Species == x)
  .spp_name <- spplookup %>% filter(Species == x) %>% pull(commonname) %>% as.character()
  
  #table of totals by length group by site
  .summ_stn <- .spp_tot %>% group_by(Station, LengthGroup) %>% 
    summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>% 
    left_join(totaleffort.df, by = c("Station" = "Station")) %>%
    mutate(CPUE_netnight = TotalCatch / Effort_netnights) %>% dplyr::select(-Effort_netnights)
  
  #table of totals by site
  .summ_stn_all <- .spp_tot %>% group_by(Station) %>% 
    summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>% 
    left_join(totaleffort.df, by = c("Station" = "Station")) %>%
    mutate(CPUE_netnight = TotalCatch / Effort_netnights) %>% dplyr::select(-Effort_netnights)
  
  #now create a summary table of totals by length group
  .summ_tbl <- .spp_tot %>% group_by(LengthGroup) %>% 
    summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>%
    mutate(CPUE_netnight = TotalCatch / totaleffort)
  # now merge these totals with mean length by length group
  .summ_tbl <- .summ_tbl %>% left_join(.spp_len %>% group_by(LengthGroup) %>% 
                                  summarise(MeanLength_mm = mean(Length, na.rm = TRUE)))
  .summ_tbl <- .summ_tbl %>% left_join(.spp_len %>% group_by(LengthGroup) %>% 
                                  summarise(MinLength_mm = min(Length, na.rm=T), 
                                            MaxLength_mm = max(Length, na.rm=T)))
  
  #create a table for all sites and length groups 
  .summ_tbl_all <- .spp_tot %>% summarise(TotalCatch = sum(totcount, na.rm = TRUE))
  .summ_tbl_all <- .summ_tbl_all %>% 
    add_column(MeanLength =  mean(.spp_len$Length, na.rm = TRUE), 
               LengthRange = as.character(paste0(min(.spp_len$Length, na.rm = TRUE), "-",  
                                                 max(.spp_len$Length, na.rm = TRUE))) )
  
  
  Ann.CPUE <- matrix(NA, curr_yr-2000, 4) # set up matrix for output, with as many columns as years
  rownames(Ann.CPUE) <- 2001:curr_yr
  colnames(Ann.CPUE) <- c("CPUE_netnights", "Catch", "Catch Rank", "CPUE Rank")
  for(i in 1:(curr_yr-2000)){
    .year <- 2000+i
    .yr.eff <- sum(as.numeric((effort$Effort[effort$Year==.year])), na.rm = T)/48  #yearly fishing effort
    .totalcatch.spp.year <- sum(allcatch$totcount[allcatch$Year==.year & allcatch$Species==x])  #total catch for year, 'spp' species
    Ann.CPUE[i,2] <- .totalcatch.spp.year  # Catch for the year for this species
    Ann.CPUE[i,1] <- .totalcatch.spp.year/.yr.eff # CPUE for the year for this species
  }
  Ann.CPUE[,3] <- rank(Ann.CPUE[,2])
  Ann.CPUE[,4] <- rank(Ann.CPUE[,1])
  
  Ann.CPUE.plot <- data.frame(Ann.CPUE[,1:2])
  Ann.CPUE.plot$Year <- as.numeric(rownames(Ann.CPUE.plot))
  
  #Make the 3 year comparison table
  threeyear_len <- all.len %>% filter(Year == curr_yr | Year == curr_yr-1 | Year == curr_yr-2, Species == x) %>% 
    group_by(Group, Year) %>% summarise(MeanLength_mm = mean(Length, na.rm = TRUE))
  
  threeyear_catch <- allcatch %>% filter(Year == curr_yr | Year == curr_yr-1 | Year == curr_yr-2, Species == x) %>% 
    group_by(LG, Year) %>% summarise(TotalCatch = sum(totcount, na.rm = TRUE))
  
  threeyear_eff <- effort %>% filter(Year == curr_yr | Year == curr_yr-1 | Year == curr_yr-2) %>% 
    group_by(Year) %>% summarise(Netnights = sum(as.numeric(Effort), na.rm=TRUE)/48)
  
  threeyearcomp <- threeyear_len %>% left_join(threeyear_catch, by = c("Year" = "Year", "Group" = "LG") ) %>% left_join(threeyear_eff, by = c("Year" = "Year")) %>%
    mutate(CPUE_netnights = TotalCatch / Netnights, Year = as.character(Year))

  # Now the printing section
  #cat('\n##', x )
  cat('\n##', .spp_name, '\n' )
  cat('\n####', curr_yr, ' - All Sites and Length Groups combined \n')
  print(kable(.summ_tbl_all, format.args = list(big.mark = ","), digits = c(0, 1, 0)) %>%
          kable_styling(full_width = FALSE))
  
  cat('\n####', curr_yr, ' - By Station & Length Group \n')
  print(kable(.summ_stn, format.args = list(big.mark = ","), digits = c(0, 0, 0, 1)) %>% 
    kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) %>%
    collapse_rows(columns = 1, valign = "top") )
  
  cat('\n####', curr_yr, ' - By Station \n')
  print(kable(.summ_stn_all, format.args = list(big.mark = ","), digits = c(0, 0, 1)) %>% 
    kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) %>%
    collapse_rows(columns = 1, valign = "top") )
  
  cat('\n####', curr_yr, ' - By Length Group, All Sites Combined \n')
  print(kable(.summ_tbl, format.args = list(big.mark = ","), digits = c(0, 0, 1, 1, 1, 1)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F))
  
  print(kable(Ann.CPUE, format.args = list(big.mark = ","), digits = c(1, 0, 0, 0)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                      full_width = F, position = "float_left")) 

  print( ggplot(Ann.CPUE.plot, aes(x = Year, y=CPUE_netnights)) + geom_line() + 
          ggtitle(paste0(.spp_name, " - CPUE by Year")))

  
  cat('\n')
  
  cat('\n#### Three Year Comparison \n')
  print(kable(threeyearcomp, format.args = list(big.mark = ","), digits = c(0, 0, 1, 0, 1, 1)) %>% 
    kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) %>%
    collapse_rows(columns = 1, valign = "top") )
  
  cat('\n')
  
  cat('\n ***** \n')
    
}



# This function is essentially a stripped down version of the above, leaving out all length calculations
spp_summ_nolen <- function(x){
  .spp_len <- all.len.year %>% filter(Species == x)
  .spp_tot <- allcatch.year %>% filter(Species == x)
  .spp_name <- spplookup %>% filter(Species == x) %>% pull(commonname) %>% as.character()
  
    #table of totals by length group by site
  .summ_stn <- .spp_tot %>% group_by(Station) %>% 
    summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>% 
    left_join(totaleffort.df, by = c("Station" = "Station")) %>%
    mutate(CPUE_netnight = TotalCatch / Effort_netnights) %>% dplyr::select(-Effort_netnights)
  
    #now create a summary table of totals by length group
  .summ_tbl <- .spp_tot %>% 
    summarise(TotalCatch = sum(totcount, na.rm = TRUE)) %>%
    mutate(CPUE_netnight = TotalCatch / totaleffort)
  
  #original code from Duncan, I just modified to transpose it. Should update this
  Ann.CPUE.nolen <- matrix(NA, curr_yr-2000, 4) # set up matrix for output, with as many columns as years
  rownames(Ann.CPUE.nolen) <- 2001:curr_yr
  colnames(Ann.CPUE.nolen) <- c("CPUE_netnights", "Catch", "Catch Rank", "CPUE Rank")
  for(i in 1:(curr_yr-2000)){
    .year <- 2000+i
    .yr.eff <- sum(as.numeric((effort$Effort[effort$Year==.year])), na.rm = T)/48  #yearly fishing effort
    .totalcatch.spp.year <- sum(allcatch$totcount[allcatch$Year==.year & allcatch$Species==x])  
    Ann.CPUE.nolen[i,2] <- .totalcatch.spp.year  # Catch for the year for this species
    Ann.CPUE.nolen[i,1] <- .totalcatch.spp.year/.yr.eff # CPUE for the year for this species
  }
  Ann.CPUE.nolen[,3] <- rank(Ann.CPUE.nolen[,2])
  Ann.CPUE.nolen[,4] <- rank(Ann.CPUE.nolen[,1])
  
  Ann.CPUE.nolen.plot <- data.frame(Ann.CPUE.nolen[,1:2])
  Ann.CPUE.nolen.plot$Year <- as.numeric(rownames(Ann.CPUE.nolen.plot))
  
  #now print the results
  cat('\n##', .spp_name, '\n' )
  cat('\n####', curr_yr, ' - By Station \n')
  print(kable(.summ_stn, format.args = list(big.mark = ","), digits = c(0, 0, 1)) %>% 
    kable_styling(bootstrap_options = c("hover", "condensed", "responsive"), full_width = F) %>%
    collapse_rows(columns = 1, valign = "top") )
  
  cat('\n####', curr_yr, ' - All Sites Combined \n')
  print(kable(.summ_tbl, format.args = list(big.mark = ","), digits = c(0, 1)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F))
  
  print(kable(Ann.CPUE.nolen, format.args = list(big.mark = ","), digits = c(1, 0, 0, 0)) %>%
        kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                      full_width = F, position = "float_left")) 
  
  print(ggplot(Ann.CPUE.nolen.plot, aes(x = Year, y=CPUE_netnights)) + geom_line() + 
          ggtitle(paste0(.spp_name, " - CPUE by Year")))
  
  cat('\n')
  cat('\n ***** \n')
  
}




```


```{r sppsummaries, echo = FALSE, warning = FALSE, results = 'asis', fig.width=5, fig.height=6.5}

spp_summ("BDWF")

spp_summ("ARCS")

spp_summ("LSCS")

spp_summ("DLVN")

spp_summ("HBWF")

spp_summ("ARCD")

spp_summ("SFCD")

spp_summ("RBSM")


spp_summ_nolen("ARFL")
spp_summ_nolen("FHSC")
# spp_summ_nolen("RBSM")

spp_summ_nolen("PINK")
#spp_summ_nolen("CHUM")



# if time, consider putting in as plotly, but this isn't necessary

```

